---
title: "RMX Estimators"
author: "Matthias Kohl"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
bibliography: rmx.bib
vignette: >
  %\VignetteIndexEntry{RMX Estimators}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{utf8}
---


## Introduction

Radius-minimax (rmx) estimators are a special class of optimally-robust statistical 
procedures [@Kohl2005; @Rieder2008]. It is a class of asymptotically linear 
estimators where infinitesimal (shrinking) neighborhoods around parametric models 
are assumed [@Rieder1994; @Ruckdeschel2006; @Kohl2010; @Kohl2012]. In addition, 
various diagnostic plots are included such as plots of influence functions, 
absolute and relative information and cniper points [@Kohl2005]. In addition
to these asymptotic optimal results we use finite-sample corrections based
on Monte-Carlo simulations [@Kohl2010a].

We provide vignettes for all models implemented, which can be viewed by

```{r, eval = FALSE}
vignette("Normal Location and Scale")
vignette("Binomial Probability")
vignette("Poisson Mean")
```


## sessionInfo
```{r}
sessionInfo()
```


## References