---
title: "A Guide to rmx"
author: "Matthias Kohl"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: true
bibliography: rmx.bib
vignette: >
  %\VignetteIndexEntry{A Guide to rmx}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{utf8}
---


## Introduction

Radius-minimax (rmx) estimators are a special class of optimally-robust statistical 
procedures [@Kohl2005; @Rieder2008]. It is a class of asymptotically linear 
estimators where infinitesimal (shrinking) neighborhoods around parametric models 
are assumed [@Rieder1994; @Ruckdeschel2006; @Kohl2010; @Kohl2012]. In addition, 
various diagnostic plots are included such as plots of influence functions, 
absolute and relative information and cniper points [@Kohl2005].


We first load the package.
```{r}
library(rmx)
```


## Normal location and scale
We first consider the estimation of mean and standard deviation (SD) of normal 
distributions. We use the chem dataset from package MASS.

```{r}
library(MASS)
data(chem)
```

We first have a look at the data by using a normal qq-plot.
```{r}
qqnorm(chem)
qqline(chem)
```

There is at least one clear outlier in the data. We compute mean, SD, median 
and MAD.

```{r}
mean(chem)
median(chem)
sd(chem)
mad(chem)
```

It's no surprise that mean and SD are strongly influenced by the clear outlier.
We now compute the rmx-estimator, where we assume between one and two outliers.

```{r}
rmx.chem <- rmx(chem, model = "norm", eps.lower = 1/24, eps.upper = 2/24)
rmx.chem
```

The results seem not be affected by the clear outlier. We take a closer look
at the result.

```{r}
coef(rmx.chem)
vcov(rmx.chem)
sqrt(diag(vcov(rmx.chem)))
bias(rmx.chem)
mse(rmx.chem)
```

As is also noted in the output, the covariance as well as the standard errors
are asymptotic values. Bias and MSE (mean squared error) correspond to the 
maximum asymptotic bias and MSE, respectively. A more comprehensive output 
containing the above values can be generated by function summary.

```{r}
summary(rmx.chem)
```

Since 

## sessionInfo
```{r}
sessionInfo()
```


## References