---
title: "Computation of finite-sample correction for normal location and scale"
author: "Matthias Kohl"
date: "26 12 2021"
output: html_document
---

# Introduction

From many simulation studies we know that our asymptotic optimally robust 
estimators tend to be too optimistic for finite sample sizes. In this
script we compute a finite-sample correction for normal location and scale 
based on Monte-Carlo simulations. That means, we replace the asymptotic radius
by a finite-sample corrected radius, which is larger or equal to the asymptotic
radius. A larger radius leads to a more conservative estimator.


# Preparations

We load the required functions and data we need for the computations.

```{r}
library(robustbase)
source("../R/ZZZinternal.R")
source("../R/rowRmxNorm.R")
load("../R/sysdata.rda")
```

We define the function that we will use for the computations.

```{r}
rowRMX <- function(x, r, k = 3L){
    mean <- robustbase::rowMedians(x, na.rm = TRUE)
    sd <- robustbase::rowMedians(abs(x-mean), na.rm = TRUE)/qnorm(0.75)
    if(r > 10){
        b <- sd*1.618128043
        const <- 1.263094656
        A2 <- b^2*(1+r^2)/(1+const)
        A1 <- const*A2
        a <- -0.6277527697*A2/sd
    }else{
        A1 <- sd^2*.getA1.norm(r)
        A2 <- sd^2*.getA2.norm(r)
        a <- sd*.geta.norm(r)
        b <- sd*.getb.norm(r)
    }
    robEst <- .kstep.norm.matrix(x = x, initial.est = cbind(mean, sd), 
                                 A1 = A1, A2 = A2, a = a, b = b, k = k,
                                 na.rm = TRUE)
    colnames(robEst$est) <- c("mean", "sd")
    robEst$est
}
```


# Least favorable contamination

We compute the contaminating distribution that leads to the maximum empirical
finite-sample MSE. 

```{r, eval = FALSE}
n <- 10
M <- 1e6
eps <- 0.01
D <- 0.1
empMSE <- function(r, x, n){
    RadMinmax <- rowRMX(x, r = r)
    n*(mean(RadMinmax[,1]^2) + mean((RadMinmax[,2]-1)^2))
}
r <- rbinom(n*M, prob = eps, size = 1)
Mid <- rnorm(n*M)
Mcont <- rep(D, n*M)
Mre <- matrix((1-r)*Mid + r*Mcont, ncol = n)
ind <- rowSums(matrix(r, ncol = n)) >= n/2
while(any(ind)){
    M1 <- sum(ind)
    cat("M1:\t", M1, "\n")
    r <- rbinom(n*M1, prob = eps, size = 1)
    Mid <- rnorm(n*M1)
    Mcont <- r(contD)(n*M1)
    Mre[ind,] <- (1-r)*Mid + r*Mcont
    ind[ind] <- rowSums(matrix(r, ncol = n)) >= n/2
}
fun1 <- function(D){
    Mcont <- rep(D, n*M)
    Mre <- matrix((1-r)*Mid + r*Mcont, ncol = n)
    empMSE(r = 1, x = Mre, n = n)
}
sapply(c(seq(0.1, 10, length = 20), 20, 50, 100, 1000, 1e4, 1e6), fun1)
```

The simulations show that a Dirac measure at some very high value leads to 
the maximum empirical MSE. We will use it as contaminating distribution in our
subsequent simulations.


# Finite-sample corrected radius

We now compute the finite-sample optimal radius for a grid of sample sizes
and contaminations. The sample size n is at least 3, as it is impossible to 
have less than 50% contamination in case of n = 2.

We define functions to determine the empirical MSE and finite-sample radius.

```{r}
empMSE <- function(r, x, n){
    RadMinmax <- rowRMX(x, r = r)
    n*(mean(RadMinmax[,1]^2) + mean((RadMinmax[,2]-1)^2))
}
finRad <- function(n, eps, D, M){
  r.fi <- numeric(length(eps))
  names(r.fi) <- eps
  i <- 0
  repeat{
    i <- i + 1
    r <- rbinom(n*M, prob = eps[i], size = 1)
    Mid <- rnorm(n*M)
    Mcont <- rep(D, n*M)
    Mre <- matrix((1-r)*Mid + r*Mcont, ncol = n)
    rm(Mid, Mcont)
    gc()
    ind <- rowSums(matrix(r, ncol = n)) >= n/2
    rm(r)
    gc()
    while(any(ind)){
      M1 <- sum(ind)
      r <- rbinom(n*M1, prob = eps[i], size = 1)
      Mid <- rnorm(n*M1)
      Mcont <- rep(D, n*M1)
      Mre[ind,] <- (1-r)*Mid + r*Mcont
      ind[ind] <- rowSums(matrix(r, ncol = n)) >= n/2
      rm(Mid, Mcont, r)
      gc()
    }
    r.fi[i] <- optimize(empMSE, 
                        interval = c(eps[i], min(max(2, n*eps[i]*25), 10)), 
                        x = Mre, n = n)$minimum
    rm(Mre)
    gc()
    if(round(r.fi[i], 2) == 1.74 | i == length(eps)) break
  }
  r.fi
}
```

To speed up the computations, we use packages foreach and doParallel.

```{r}
library(foreach)
library(doParallel)
M <- 1e2
D <- 1e6
n <- c(3:50, seq(55, 100, by = 5), seq(110, 200, by = 10), seq(250, 500, by = 25))
eps <- c(seq(0.001, 0.05, by = 0.001), seq(0.06, to = 0.5, by = 0.01))
cores <- detectCores()
cl <- makeCluster(cores[1]-1)
registerDoParallel(cl)
```

We compute the finite-sample radii.

```{r}
r.fi <- foreach(i=seq(along = n)) %dopar% {
  finRad(n[i], eps, D, M)
}
```

We take a look at the results.

```{r}
r.fi <- do.call("cbind", r.fi)
save(r.fi, file = "FiniteSampleRadiusNorm.RData")
r.as <- outer(eps, sqrt(n))
## no convergence resp. wrong optima
sum(r.fi > 1.744, na.rm = TRUE)
r.fi[r.fi > 1.744 & !is.na(r.fi)]
```

```{r}
r.finit <- r.fi
## setze auf NA
r.finit[r.fi > 1.744 & !is.na(r.fi)] <- NA
r.finit[r.finit > 1.744 & !is.na(r.finit)]
sum(r.finit > r.as & !is.na(r.finit), na.rm = TRUE)
sum(!is.na(r.finit))

## r.finit muss größer oder gleich r.as sein
r.finit[r.finit <= r.as & !is.na(r.finit)]
r.as[r.finit <= r.as & !is.na(r.finit)]
r.finit[r.finit <= r.as & !is.na(r.finit)] <- r.as[r.finit <= r.as & !is.na(r.finit)]
r.finit[r.finit < r.as & !is.na(r.finit)]

for(i in 1:nrow(r.finit)){
    cat("====================================================================\n")
    cat(eps[i], "\n")
    ind <- r.finit[i,] > 1.744
    fit <- lm(r.finit[i,!ind] ~ I(1/sqrt(n[!ind])) + I(1/n[!ind]) + I(1/n[!ind]^(3/2)))
    print(summary(fit))
#    fit <- lm(r.finit[i,!ind] ~ I(1/n[!ind]^(3/2)))
#    print(summary(fit))
}

## ab ca. 1.655 keine Änderung mehr sichtbar
r.finit[18,]
r.finit2 <- r.finit
## minimales b, ab ca. r = 1.67
r.finit2[is.na(r.finit) | r.finit > 1.655] <- 1.67

plot(n, r.finit[19,])
plot(n, r.finit2[19,])

for(i in c(1:11)){
    r.finit2[i, r.finit2[i,] > 1.655] <- approx(n[r.finit2[i,] <= 1.655], 
                                                r.finit2[i,r.finit2[i,] <= 1.655], 
                                                n[r.finit2[i,] > 1.655], rule = 2)$y
}
r.finit2[12,2] <- 1.67

r.finit2 <- round(r.finit2, 4)

#load("sysdata.rda")
#.fsRadius.norm <- r.finit2
#save(.fsRadius.norm, .radius.gitter.norm, .A1.norm, .A2.norm, 
#     .a.norm, .b.norm, file = "sysdata.rda")
```
